---
output:
  pdf_document: default
  html_document: default
---
Installing required packages
```{r}
#Delete the # on Line 4 to allow the code to work
#You only need to install packages once
#install.packages(c("metafor","meta"))
```
Loading packages
```{r}
library(metafor)
library(meta)
```
Reading in data
```{r}
#We use the same data set for the logit and double arcsine transformations 
dat=read.csv("./data.csv",header=T,sep=",")
```
Calculating overall summary proportion
```{r}
#If you want to use the harmonic mean, delete the # on Line 18 and add a # on Line 19
ies.da=escalc(xi=cases,ni=total,measure="PFT",data=dat,add=0)
pes.da=rma(yi,vi,data=ies.da,method="DL",weighted=TRUE)
#pes=predict(pes.da,transf=transf.ipft.hm,targ=list(ni=dat$total)) #Use harmonic mean
pes=predict(pes.da,transf=transf.ipft.hm,targ = list(ni=1/(pes.da$se)^2))
print(pes,digits=6);print(pes.da,digits=6);confint(pes.da,digits=2)
```
Identifying outliers with residuals
```{r}
stud.res=rstudent(pes.da)
abs.z=abs(stud.res$z)
stud.res[order(-abs.z)]
```
Leave-one-out analysis
```{r}
#I commented out the code pieces that use harmonic mean
l1o=leave1out(pes.da)
yi=l1o$estimate;vi=l1o$se^2
#forest(yi,vi,transf=transf.ipft.hm,targ=list(ni=dat$total),slab=paste(dat$author,dat$year,sep=", "),xlab="Summary proportions leaving out each study",refline=pes$pred,digits=6)
forest(yi,vi,transf=transf.ipft.hm,targ = list(ni=1/(pes.da$se)^2),slab=paste(dat$author,dat$year,sep=", "),xlab="Summary proportions leaving out each study",refline=pes$pred,digits=6)
#L1O=leave1out(pes.da,transf=transf.ipft.hm,targ=list(ni=dat$total));print(L1O,digits=6)
L1O=leave1out(pes.da,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2));print(L1O,digits=6)
```
Baujat plot
```{r}
baujat(pes.da)
```
Diagnostic tests
```{r}
inf=influence(pes.da)
print(inf);plot(inf)
```
Removing outliers
```{r}
ies.da.noutlier=escalc(xi=cases,ni=total,measure="PFT", data=dat[-c(11,18),],add=0)
pes.da.noutlier=rma(yi,vi,data=ies.da.noutlier,method="DL",weighted=TRUE)
#pes.noutlier=predict(pes.da.noutlier,transf=transf.ipft.hm,targ=list(ni=dat$total))
pes.noutlier=predict(pes.da.noutlier,transf=transf.ipft.hm,targ=list(ni=1/(pes.da.noutlier$se)^2))
#print(pes,digits=4)
print(pes.noutlier,digits=4)
```
Calculating subgroup summary proportions, conducting subgroup analysis, and recalculating summary proportion assuming a common between-study variance component across subgroups
```{r}
# For those who prefer using the double arcsine transformation, we recommend constructing forest plots with metafor, and here’s why: When you apply the double arcsine transformation, it’s important to note that the metaprop() function employs Eq.10 for back-transformation, as opposed to Eq.13 recommended by Barendregt et al. (2013).  The rma() and predict() functions in metafor have the flexibility to use either Eq.10 or Eq.13, based on the user's preference. This distinction means that proportions estimated by metafor using Eq.13 will not align with those calculated through the metaprop() function. Consequently, if one begins with metafor for estimation and back-transformation of proportions and then switches to the meta package for the generation of a forest plot, you’ll encounter inconsistencies in the estimates of proportions. To avoid such an issue, we suggest readers consistently use metafor throughout the process when you use the double arcsine transformation.
subganal.studydesign=rma(yi,vi,data=ies.da,mods=~studydesign,method="DL")
pes.da.birthcohort=rma(yi,vi,data=ies.da,mods=~studydesign=="Others",method="DL")
pes.da.others=rma(yi,vi,data=ies.da,mods=~studydesign=="Birth cohort",method="DL")
#pes.subg.studydesign=predict(subganal.studydesign,transf=transf.ipft.hm,targ=list(ni=dat$total))
pes.subg.studydesign=predict(subganal.studydesign,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
dat.samevar=data.frame(estimate=c((pes.da.birthcohort$b)[1],(pes.da.others$b)[1]),
                       stderror=c((pes.da.birthcohort$se)[1],(pes.da.others$se)[1]),
                       tau2=subganal.studydesign$tau2)
pes.da.studydesign=rma(estimate,sei=stderror,method="FE",data=dat.samevar)
#pes.studydesign=predict(pes.da.studydesign,transf=transf.ipft.hm,targ=list(ni=dat$total))
pes.studydesign=predict(pes.da.studydesign,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
#print(pes.subg.studydesign[order(pes.subg.studydesign$pred)],digits=6)
print(pes.studydesign,digits=6)
print(pes.subg.studydesign[1],digits=6)
print(pes.subg.studydesign[17],digits=6)
print(subganal.studydesign,digits=4)
```
Calculating subgroup summary proportions, conducting subgroup analysis, and recalculating summary proportion not assuming a common between-study variance component across subgroups
```{r}
pes.da.birthcohort=rma(yi,vi,data=ies.da,subset=studydesign=="Birth cohort",method="DL")
pes.da.others=rma(yi,vi,data=ies.da,subset=studydesign=="Others",method="DL")
#pes.birthcohort=predict(pes.da.birthcohort,transf=transf.ipft.hm,targ=list(ni=dat$total))
#pes.others=predict(pes.da.others,transf=transf.ipft.hm,targ=list(ni=dat$total))
pes.birthcohort=predict(pes.da.birthcohort,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
pes.others=predict(pes.da.others,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
dat.diffvar= data.frame(estimate=c(pes.da.birthcohort$b, pes.da.others$b),
                        stderror=c(pes.da.birthcohort$se, pes.da.others$se),
                        studydesign = c("Birth cohort","Others"), 
                        tau2=round(c(pes.da.birthcohort$tau2, pes.da.others$tau2),3))
subganal.studydesign=rma(estimate, sei=stderror, mods=~studydesign, method="FE", data=dat.diffvar, digits=4)
pes.da.studydesign=rma(estimate, sei=stderror, method="FE", data=dat.diffvar, digits=4) 
#pes.studydesign=predict(pes.da.studydesign,transf=transf.ipft.hm,targ=list(ni=dat$total))
pes.studydesign=predict(pes.da.studydesign,transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
print(pes.birthcohort,digits=6);print(pes.da.birthcohort,digits=2)
print(pes.others,digits=6);print(pes.da.others,digits=2)
print(subganal.studydesign,digits=2)
print(pes.studydesign,digits=6)

```
Forest plot with metafor assuming a common variance component
```{r}
png("forestplotmetafor.png",width=1000,height=1000)
ies.summary=summary(ies.da,transf=transf.ipft,ni=dat$total)
forest(ies.summary$yi,
       order = ies.summary$studesg,
       ci.lb=ies.summary$ci.lb,ci.ub=ies.summary$ci.ub,
       slab=paste(dat$author,dat$year, sep=", "),
       ilab=cbind(data=dat$cases, dat$total),
       ilab.xpos=c(-0.002,-0.001),
       rows=c(19:14, 8.5:-1.5),
       at=c(seq(from=0,to=0.004,by=0.001)), 
       refline=pes.studydesign$pred,
       ylim=c(-5,23),
       xlim=c(-0.005,0.005),
       bg="navy",
       cex=1,
       main="Effect Size by Study Design (Mixed-Effects Model)",
       xlab="Proportion",
       digits=6
       )
addpoly(pes.subg.studydesign[1]$pred,ci.lb=pes.subg.studydesign[1]$ci.lb,ci.ub=pes.subg.studydesign[1]$ci.ub,row=12.7,col="red",digits=5,cex=1.1)
addpoly(pes.subg.studydesign[17]$pred,ci.lb=pes.subg.studydesign[17]$ci.lb,ci.ub=pes.subg.studydesign[17]$ci.ub,row=-2.7,col="red",digits=5,cex=1.1)
addpoly(pes.studydesign$pred,ci.lb=pes.studydesign$ci.lb,ci.ub=pes.studydesign$ci.ub,row=-4.4,col="red",digits=5,cex=1.1)
text(-0.005,21.8,pos=4,cex=1.1,"Author, Year")
text(c(-0.0026,-0.0014),21.8,pos=4,cex=1.1,c("Cases", "Total"))
text(0.0023,21.8,pos=4,cex=1.1,font=1,"Proportion [95% CI]")
text(-0.005,c(9.7,20.2),pos=4, cex=1.1,c("Others", "Birth cohort"))
text(-0.005,-4.4,pos=4,cex=1.1,bquote(paste("Overall")))
text(-0.005,12.7,pos=4,cex=1.1,bquote(paste("Subgroup")))
text(-0.005,-2.7,pos=4,cex=1.1,bquote(paste("Subgroup")))
abline(h=-3.5)
dev.off()
```
Forest plot with metafor not assuming a common variance component
```{r}
png("forestplotmetafor.png",width=550,height=550)
ies.summary=summary(ies.da,transf=transf.ipft,ni=dat$total)
par(cex=1,font=6)
forest(ies.summary$yi,
       order = ies.summary$studesg,
       ci.lb=ies.summary$ci.lb,ci.ub=ies.summary$ci.ub,
       ylim=c(-5,23),
       xlim=c(-0.005,0.005),
       slab=paste(dat$author,dat$year, sep=", "),
       ilab=cbind(data=dat$cases, dat$total),
       ilab.xpos=c(-0.0019,-0.0005),
       ilab.pos=2,
       rows=c(19:14, 8.5:-1.5),
       at=c(seq(from=0,to=0.004,by=0.001)), 
       refline=pes.studydesign$pred,
       main="",
       xlab="Proportion (%)",
       digits=4
       )
par(cex=1.2,font=7)
addpoly(pes.birthcohort$pred,ci.lb=pes.birthcohort$ci.lb,ci.ub=pes.birthcohort$ci.ub,row=12.8,digits=6)
addpoly(pes.others$pred,ci.lb=pes.others$ci.lb,ci.ub=pes.others$ci.ub,row=-2.7,digits=6)
addpoly(pes.studydesign$pred,ci.lb=pes.studydesign$ci.lb,ci.ub=pes.studydesign$ci.ub,row=-4.6,digits=6)
par(cex=1.1,font=7)
text(-0.005,21.8,pos=4,"Study")
text(c(-0.0026,-0.0014),21.8,pos=4,c("Cases", "Total"))
text(0.0025,21.8,pos=4,"Proportion [95% CI]")
text(-0.005,c(9.7,20.2),pos=4,c("Others", "Birth cohort"))
par(cex=1,font=7)
text(-0.005,-4.6,pos=4,c("Overall"))
text(-0.005,12.8,pos=4,c("Subgroup"))
text(-0.005,-2.7,pos=4,c("Subgroup"))
abline(h=-3.7)
dev.off()
```
Forest plot with meta assuming a common between variance
```{r}
subganal.studydesign=rma(yi,vi,data=ies.da,mods=~studydesign,method="DL")
pes.summary=metaprop(cases,
                     total,
                     author,
                     data=dat,
                     byvar=studydesign,
                     sm="PFT",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=TRUE,
                     tau.preset=sqrt(subganal.studydesign$tau2),
                     title="")
png("forestplotmeta.png",width=1000,height=1000)
forest(pes.summary,
       xlim = c(0,1), pscale =1, 
       rightcols=c("effect", "ci","w.random"),
       rightlabs=c("Proportion", "95% C.I.","Weights"),
       leftcols = c("studlab", "event", "n"), 
       leftlabs = c("Study", "Cases", "Total"),
       xlab = "Prevalence", 
       fs.xlab=12,
       fs.study=12,
       fs.study.lables=12,
       fs.heading=12,
       squaresize = 0.5, col.square="navy", col.square.lines="navy",
       col.diamond="navy", col.diamond.lines="navy",
       comb.fixed=FALSE,
       comb.random=TRUE,
       resid.hetstat = FALSE,
       lty.fixed=0,
       lty.random=2, 
       type.study="square",
       type.random="diamond",
       ff.fixed="bold.italic",
       ff.random="bold.italic",
       hetlab = "Heterogeneity:",
       fs.hetstat=10,
       smlab="",
       print.Q=TRUE,
       print.pval.Q=TRUE,
       print.I2=TRUE,
       print.tau2=TRUE,
       col.by="grey",
       digits=6)
dev.off()
```
Scatterplot for study design
```{r}
#Using R's native functions
subganal.studesg=rma(yi,vi,data=ies.da,mods=~studesg,method="DL")
#preds.studesg=predict(subganal.studesg,newmods=c(0:2),transf=transf.ipft.hm,targ=list(ni=dat$total))
preds.studesg=predict(subganal.studesg,newmods=c(0:2),transf=transf.ipft.hm,targ=list(ni=1/(pes.da$se)^2))
wi=1/sqrt(ies.da$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
#plot(ies.da$studesg,transf.ipft.hm(ies.da$yi,targ=list(ni=dat$total)),cex=size,pch=1,las=1,xlab="Study design", ylab="Proportions")
plot(ies.da$studesg,transf.ipft.hm(ies.da$yi,targ=list(ni=1/(pes.da$se)^2)),cex=size,pch=1,las=1,xlab="Study design", ylab="Proportions")
lines(0:2,preds.studesg$pred,col="navy")
lines(0:2,preds.studesg$ci.lb,lty="dashed",col="maroon")
lines(0:2,preds.studesg$ci.ub,lty="dashed",col="maroon")
ids=c(1:6)
pos=c(3)
#text(ies.da$studesg[ids],transf.ipft.hm(ies.da$yi,targ=list(ni=dat$total))[ids],ids,cex=0.9,pos=pos)
text(ies.da$studesg[ids],transf.ipft.hm(ies.da$yi,targ=list(ni=1/(pes.da$se)^2))[ids],ids,cex=0.9,pos=pos)
```
Scatterplot for study design using regplot()
```{r}
subganal.studesg=rma(yi,vi,data=ies.da,mods=~studesg,method="DL")
regplot(subganal.studesg, mod = "studesg",
        xlab="Study design",
        transf=transf.ipft.hm,
        #targ=list(ni=dat$total),
        targ=list(ni=1/(pes.da$se)^2),
        legend = "topleft",
        label = TRUE,
        shade = "white",
        bg = "white",
        lcol = "navy",
        digits = 6)
```