---
output:
  pdf_document: default
  html_document: default
---
Installing required packages
```{r}
#Delete the # on Line 4 to allow the code to work
#You only need to install packages once
#install.packages(c("metafor","meta"))
```
Loading packages
```{r}
library(metafor)
library(meta)
```
Reading in data
```{r}
#We use the same data set for the logit and double arcsine transformations 
dat=read.csv("./data.csv",header=T,sep=",")
```
Calculating overall summary proportion
```{r}
ies.logit=escalc(xi=cases,ni=total,measure="PLO",data=dat)
pes.logit=rma(yi,vi,data=ies.logit,method="DL",weighted=TRUE)
pes=predict(pes.logit,transf=transf.ilogit)
print(pes,digits=6);print(pes.logit,digits=4);confint(pes.logit,digits=2)
```
Identifying outliers with residuals
```{r}
stud.res=rstudent(pes.logit)
abs.z=abs(stud.res$z)
stud.res[order(-abs.z)]
```
Leave-one-out analysis
```{r}
l1o=leave1out(pes.logit)
yi=l1o$estimate;vi=l1o$se^2
forest(yi,vi,transf=transf.ilogit,slab=paste(dat$author,dat$year,sep=", "),xlab="Summary proportions leaving out each study",refline=pes$pred,digits=6)
abline(h=0.1)
L1O=leave1out(pes.logit,transf=transf.ilogit);print(L1O,digits=6)
```
Baujat plot
```{r}
bjplot <- baujat(pes.logit,
                 symbol=19, 
                 xlim=c(0,15),
                 xlab="Contribution to Overall Heterogeneity", 
                 ylab="Influence on Summary Proportion")
bjplot <- bjplot[bjplot$x >= 10 | bjplot$y >= 0.4,]
text(bjplot$x, bjplot$y, bjplot$slab, pos=1)
```
Diagnostic tests
```{r}
inf=influence(pes.logit)
print(inf);plot(inf)
```
Removing outliers
```{r}
ies.logit.noutlier=escalc(xi=cases,ni=total,measure="PLO", data=dat[-c(2,8),])
pes.logit.noutlier=rma(yi,vi,data=ies.logit.noutlier,method="DL")
pes.noutlier=predict(pes.logit.noutlier,transf=transf.ilogit)
print(pes.noutlier,digits=6)
```
Calculating subgroup summary proportions, conducting subgroup analysis, and recalculating summary proportion assuming a common between-study variance component across subgroups
```{r}
subganal.studydesign=rma(yi,vi,data=ies.logit,mods=~studydesign,method="DL")
pes.logit.birthcohort=rma(yi,vi,data=ies.logit,mods=~studydesign=="Others",method="DL")
pes.logit.others=rma(yi,vi,data=ies.logit,mods=~studydesign=="Birth cohort",method="DL")
pes.subg.studydesign=predict(subganal.studydesign,transf=transf.ilogit)
dat.samevar=data.frame(estimate=c((pes.logit.birthcohort$b)[1],(pes.logit.others$b)[1]),
                       stderror=c((pes.logit.birthcohort$se)[1],(pes.logit.others$se)[1]),
                       tau2=subganal.studydesign$tau2)
pes.logit.studydesign=rma(estimate,sei=stderror,method="FE",data=dat.samevar) 
pes.studydesign=predict(pes.logit.studydesign,transf=transf.ilogit)
#print(pes.subg.studydesign[order(pes.subg.studydesign$pred)],digits=6)
print(subganal.studydesign,digits=4)
print(pes.subg.studydesign[1],digits=6);print(pes.subg.studydesign[17],digits=6)
print(pes.studydesign,digits=6)
```
Calculating subgroup summary proportions, conducting subgroup analysis, and recalculating summary proportion not assuming a common between-study variance component across subgroups
```{r}
pes.logit.birthcohort=rma(yi,vi,data=ies.logit,subset=studydesign=="Birth cohort",method="DL")
pes.logit.others=rma(yi,vi,data=ies.logit,subset=studydesign=="Others",method="DL")
pes.birthcohort=predict(pes.logit.birthcohort,transf=transf.ilogit,digits=5)
pes.others=predict(pes.logit.others,transf=transf.ilogit,digits=5)
dat.diffvar=data.frame(estimate=c(pes.logit.birthcohort$b,pes.logit.others$b),
                       stderror=c(pes.logit.birthcohort$se,pes.logit.others$se),
                       studydesign=c("Birth cohort","Others"), 
                       tau2=round(c(pes.logit.birthcohort$tau2, pes.logit.others$tau2),3))
subganal.studydesign=rma(estimate,sei=stderror,mods=~studydesign,method="FE",data=dat.diffvar)
pes.logit.studydesign=rma(estimate,sei=stderror,method="FE",data=dat.diffvar)
pes.studydesign=predict(pes.logit.studydesign,transf=transf.ilogit)
print(pes.birthcohort,digits=6);print(pes.logit.birthcohort,digits=3)
print(pes.others,digits=6);print(pes.logit.others,digits=3)
print(subganal.studydesign,digits=3)
print(pes.studydesign,digits=6)
```
Forest plot with metafor assuming a common variance component
```{r}
ies.summary=summary(ies.logit,transf=transf.ilogit)
png("forestplotmetafor.png",width=550,height=560)
forest(ies.summary$yi,
       order = ies.summary$studesg,
       ci.lb=ies.summary$ci.lb,ci.ub=ies.summary$ci.ub,
       slab=paste(dat$author,dat$year, sep=", "),
       ilab=cbind(data=dat$cases, dat$total),
       ilab.xpos=c(-0.002,-0.001),
       rows=c(19:14, 8.5:-1.5),
       at=c(seq(from=0,to=0.004,by=0.001)), 
       refline=pes.studydesign$pred,
       ylim=c(-5,23),
       xlim=c(-0.005,0.005),
       bg="navy",
       cex=1,
       main="Effect Size by Study Design (Mixed-Effects Model)",
       xlab="Proportion",
       digits=4
       )
addpoly(pes.subg.studydesign[1]$pred,ci.lb=pes.subg.studydesign[1]$ci.lb,ci.ub=pes.subg.studydesign[1]$ci.ub,row=12.7,col="red",digits=5,cex=1.1)
addpoly(pes.subg.studydesign[17]$pred,ci.lb=pes.subg.studydesign[17]$ci.lb,ci.ub=pes.subg.studydesign[17]$ci.ub,row=-2.7,col="red",digits=5,cex=1.1)
addpoly(pes.studydesign$pred,ci.lb=pes.studydesign$ci.lb,ci.ub=pes.studydesign$ci.ub,row=-4.4,col="red",digits=5,cex=1.1)
text(-0.005,21.8,pos=4,cex=1.1,"Author, Year")
text(c(-0.0026,-0.0014),21.8,pos=4,cex=1.1,c("Cases", "Total"))
text(0.0023,21.8,pos=4,cex=1.1,font=1,"Proportion [95% CI]")
text(-0.005,c(9.7,20.2),pos=4, cex=1.1,c("Others", "Birth cohort"))
text(-0.005,-4.4,pos=4,cex=1.1,bquote(paste("Overall")))
text(-0.005,12.7,pos=4,cex=1.1,bquote(paste("Subgroup")))
text(-0.005,-2.7,pos=4,cex=1.1,bquote(paste("Subgroup")))
abline(h=-3.5)
dev.off()
```
Forest plot with metafor not assuming a common variance component
```{r}
ies.summary=summary(ies.logit,transf=transf.ilogit)
png("forestplotmetafor.png",width=550,height=550)
par(cex=1,font=6)
forest(ies.summary$yi,
       order = ies.summary$studesg,
       ci.lb=ies.summary$ci.lb,ci.ub=ies.summary$ci.ub,
       ylim=c(-5,23),
       xlim=c(-0.005,0.005),
       slab=paste(dat$author,dat$year, sep=", "),
       ilab=cbind(data=dat$cases, dat$total),
       ilab.xpos=c(-0.0019,-0.0005),
       ilab.pos=2,
       rows=c(19:14, 8.5:-1.5),
       at=c(seq(from=0,to=0.004,by=0.001)), 
       refline=pes.studydesign$pred,
       main="",
       xlab="Proportion (%)",
       digits=4
       )
par(cex=1.2,font=7)
addpoly(pes.birthcohort$pred,ci.lb=pes.birthcohort$ci.lb,ci.ub=pes.birthcohort$ci.ub,row=12.8,digits=5)
addpoly(pes.others$pred,ci.lb=pes.others$ci.lb,ci.ub=pes.others$ci.ub,row=-2.7,digits=5)
addpoly(pes.studydesign$pred,ci.lb=pes.studydesign$ci.lb,ci.ub=pes.studydesign$ci.ub,row=-4.6,digits=5)
par(cex=1.1,font=7)
text(-0.005,21.8,pos=4,"Study")
text(c(-0.0026,-0.0014),21.8,pos=4,c("Cases", "Total"))
text(0.0025,21.8,pos=4,"Proportion [95% CI]")
text(-0.005,c(9.7,20.2),pos=4,c("Others", "Birth cohort"))
par(cex=1,font=7)
text(-0.005,-4.6,pos=4,c("Overall"))
text(-0.005,12.8,pos=4,c("Subgroup"))
text(-0.005,-2.7,pos=4,c("Subgroup"))
abline(h=-3.7)
dev.off()
```
Forest plot with meta assuming a common variance component
```{r}
#precision=1/(ies.logit$vi)
subganal.studydesign=rma(yi,vi,data=ies.logit,mods=~studydesign,method="DL")
pes.summary=metaprop(cases,
                     total,
                     authoryear,
                     data=dat,
                     #sortvar(precision)
                     byvar=studydesign,
                     sm="PLO",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=TRUE,
                     tau.preset=sqrt(subganal.studydesign$tau2),
                     title="")
png("forestplotmeta.png",width=700,height=600)
precison=sqrt(ies.logit$vi)
forest(pes.summary,
       #sortvar=precision,
       xlim = c(0,4), pscale =1000, 
       rightcols=FALSE,
       resid.hetstat = FALSE,
       leftcols = c("studlab","effect", "ci"),
       leftlabs = c("Study","Proportion", "95% C.I."),
       xlab = "Proportion (‰)", 
       fs.xlab=12,
       fs.study=12,
       fs.study.lables=12,
       fs.heading=12,
       squaresize = 0.5, col.square="navy", col.square.lines="navy",
       col.diamond="maroon", col.diamond.lines="maroon",
       comb.fixed=FALSE,
       lty.random=2, #lty.fixed=0,
       type.study="square",
       type.random="diamond",
       ff.random="bold", #ff.fixed="bold",
       hetlab = "Heterogeneity:",
       fs.hetstat=10,
       smlab="",
       print.Q=TRUE,
       print.pval.Q=TRUE,
       print.I2=TRUE,
       print.tau2=TRUE,
       col.by="black",
       digits=4,
       digits.Q=0)
dev.off()
```
Scatterplot for study design
```{r}
#Using R's native functions
subganal.studesg=rma(yi,vi,data=ies.logit,mods=~studesg,method="DL")
preds.studesg=predict(subganal.studesg,newmods=c(0:2),transf=transf.ilogit)
wi=1/sqrt(ies.logit$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(ies.logit$studesg,transf.ilogit(ies.logit$yi),cex=size,pch=1,las=0,xlab="Study Design", ylab="Proportion")
lines(0:2,preds.studesg$pred,col="navy")
lines(0:2,preds.studesg$ci.lb,lty="dashed",col="maroon")
lines(0:2,preds.studesg$ci.ub,lty="dashed",col="maroon")
ids=c(1:6)
pos=c(2)
text(ies.logit$studesg[ids],transf.ilogit(ies.logit$yi)[ids],ids,cex=0.9,pos=pos)
```
Scatterplot for sample size
```{r}
#Using R's native functions
subganal.size=rma(yi,vi,data=ies.logit,mods=~size,method="DL")
preds.size=predict(subganal.size,newmods=c(0:2),transf=transf.ilogit)
wi=1/sqrt(ies.logit$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
plot(ies.logit$size,transf.ilogit(ies.logit$yi),cex=size,xlab="Sample size", ylab="Proportion")
lines(0:2,preds.size$pred,col="navy")
lines(0:2,preds.size$ci.lb,lty="dashed",col="maroon")
lines(0:2,preds.size$ci.ub,lty="dashed",col="maroon")
ids=c(1:6)
pos=c(2)
text(ies.logit$studesg[ids],transf.ilogit(ies.logit$yi)[ids],ids,cex=0.9,pos=pos)
```
Scatterplot for year
```{r}
#Using R's native functions
metareg.year=rma(yi,vi,data=ies.logit,mods=~year,method="DL")
wi=1/sqrt(ies.logit$vi)
size=1+3*(wi-min(wi))/(max(wi)-min(wi))
preds.year=predict(metareg.year,newmods=c(1985:2020),transf=transf.ilogit)
plot(ies.logit$year,transf.ilogit(ies.logit$yi),cex=size,pch=1,xlab="Publication year", ylab="Proportion")
lines(1985:2020,preds.year$pred,col="navy")
lines(1985:2020,preds.year$ci.lb,lty="dashed",col="maroon")
lines(1985:2020,preds.year$ci.ub,lty="dashed",col="maroon")
ids=c(1:17)
pos=c(3)
text(ies.logit$year[ids],transf.ilogit(ies.logit$yi)[ids],ids,cex=0.9,pos=pos)
```
Scatterplot for study design using regplot()
```{r}
#Straight regression and 95%CI lines
subganal.studesg=rma(yi,vi,data=ies.logit,mods=~studesg,method="DL")
regplot(subganal.studesg, mod = "studesg", 
        xlab = "Study Design", 
        transf=transf.ilogit, 
        legend = FALSE,
        label = TRUE,
         pred=FALSE, 
        ci=FALSE,
        shade = "white",
        bg = "transparent",
        lcol = "navy",
        digits = 4)
preds.studesg=predict(subganal.studesg,newmods=c(0:1),transf=transf.ilogit)
lines(0:1,preds.studesg$pred,col="navy")
lines(0:1,preds.studesg$ci.lb,lty="dashed",col="maroon")
lines(0:1,preds.studesg$ci.ub,lty="dashed",col="maroon")
```
Scatterplot for sample size using regplot()
```{r}
subganal.size=rma(yi,vi,data=ies.logit,mods=~size,method="DL")
regplot(subganal.size, mod = "size", 
        xlab = "Sample size", 
        transf=transf.ilogit, 
        legend = "topright",
        label = TRUE,
        shade = "white",
        bg = "transparent",
        lcol = "navy",
        digits = 4)
```
Scatterplot for year using regplot()
```{r}
metareg.year=rma(yi,vi,data=ies.logit,mods=~year,method="DL")
regplot(metareg.year, 
        mod = "year", 
        transf=transf.ilogit,
        xlab = "Year of publication",
        legend = "topleft",
        label = TRUE,
        shade = "white",
        bg = "transparent",
        lcol = "navy",
        digits = 6)
```